### 2021-04-14 priority-gv.tex
Rewrote intro.
Consistently reference extended versions.

\subsubsection*{Operational Semantics}
Priority GV terms are evaluated as part of a configuration of processes. Configurations are defined by the following grammar:
\[
  \tm{\phi}
  \Coloneqq \tm{\main}
  \sep      \tm{\child}
  \hfill
  \tm{\conf{C}}, \tm{\conf{D}}, \tm{\conf{E}}
  \Coloneqq \tm{\phi\;M}
  \sep      \tm{\ppar{\conf{C}}{\conf{D}}}
  \sep      \tm{\res{x}{x'}{\conf{C}}}
  \hfill
\]
Configurations ($\tm{\conf{C}}$, $\tm{\conf{D}}$, $\tm{\conf{E}}$) consist of threads $\tm{\phi\;M}$, parallel compositions $\tm{\ppar{\conf{C}}{\conf{D}}}$, and name restrictions $\tm{\res{x}{x'}{\conf{C}}}$. To preserve the functional nature of PGV, where programs return a single value, we use flags ($\tm{\phi}$) to differentiate between the main thread, marked $\tm{\main}$, and child threads created by $\tm{\spawn}$, marked $\tm{\child}$. Only the main thread returns a value. We determine the flag of a configuration by combining the flags of all threads in that configuration:
\[
  \tm{\main}  + \tm{\child} = \tm{\main}
  \hfill
  \tm{\child} + \tm{\main}  = \tm{\main}
  \hfill
  \tm{\child} + \tm{\child} = \tm{\child}
  \hfill
  (\tm{\main}  + \tm{\main} \; \text{is undefined})
  \hfill
\]
The use of $\tm{\child}$ for child threads~\cite{lindleymorris15} overlaps with the use of the meta-variable $\cs{o}$ for priorities~\cite{dardhagay18}. Both are used to annotate sequents: flags appear on the sequent in configuration typing, and priorities in term typing. To distinguish the two symbols, they are typeset in a different font and a different colour.

@@

The only non-standard feature of the typing rules is the priority annotations. Priorities are based on \emph{obligations/capabilities} used by Kobayashi~\cite{kobayashi06}, and simplified to single priorities following Padovani~\cite{padovani14}. The integration of priorities into GV is adapted from Padovani and Novara~\cite{padovaninovara15}. Paraphrasing Dardha and Gay~\cite{dardhagay18}, priorities obey the following two laws:
\begin{enumerate*}[label= (\roman*) ]
\item an action with lower priority happens before an action with higher priority; and
\item communication requires \emph{equal} priorities for dual actions.
\end{enumerate*}

@@

Progress for the term language is standard for GV, and deviates from progress for linear $\lambda$-calculus only in that terms may reduce to values or \emph{ready terms}:
\begin{restatablelemma}{lempgvopenprogressterms}[Progress, $\tred$]%
  \label{lem:pgv-open-progress-terms}
  If $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$ and $\ty{\Gamma}$ contains only session types, then:
  \begin{enumerate*}[label=\emph{(\alph*)}]
  \item $\tm{M}$ is a value;
  \item $\tm{M}\tred\tm{N}$ for some $\tm{N}$; or
  \item $\tm{M}$ is ready.
  \end{enumerate*}
\end{restatablelemma}

>>>>

\subsubsection*{Operational Semantics}
Priority GV terms are evaluated as part of a configuration of processes. Configurations are defined by the following grammar:
\[
  \tm{\phi}
  \Coloneqq \tm{\main}
  \sep      \tm{\child}
  \hfill
  \tm{\conf{C}}, \tm{\conf{D}}, \tm{\conf{E}}
  \Coloneqq \tm{\phi\;M}
  \sep      \tm{\ppar{\conf{C}}{\conf{D}}}
  \sep      \tm{\res{x}{x'}{\conf{C}}}
  \hfill
\]
Configurations ($\tm{\conf{C}}$, $\tm{\conf{D}}$, $\tm{\conf{E}}$) consist of threads $\tm{\phi\;M}$, parallel compositions $\tm{\ppar{\conf{C}}{\conf{D}}}$, and name restrictions $\tm{\res{x}{x'}{\conf{C}}}$. To preserve the functional nature of PGV, where programs return a single value, we use flags ($\tm{\phi}$) to differentiate between the main thread, marked $\tm{\main}$, and child threads created by $\tm{\spawn}$, marked $\tm{\child}$. Only the main thread returns a value. We determine the flag of a configuration by combining the flags of all threads in that configuration:
\[
  \tm{\main}  + \tm{\child} = \tm{\main}
  \hfill
  \tm{\child} + \tm{\main}  = \tm{\main}
  \hfill
  \tm{\child} + \tm{\child} = \tm{\child}
  \hfill
  (\tm{\main}  + \tm{\main} \; \text{is undefined})
  \hfill
\]
The use of $\tm{\child}$ for child threads~\cite{lindleymorris15} overlaps with the use of the meta-variable $\cs{o}$ for priorities~\cite{dardhagay18extended}. Both are used to annotate sequents: flags appear on the sequent in configuration typing, and priorities in term typing. To distinguish the two symbols, they are typeset in a different font and a different colour.

@@

The only non-standard feature of the typing rules is the priority annotations. Priorities are based on \emph{obligations/capabilities} used by Kobayashi~\cite{kobayashi06}, and simplified to single priorities following Padovani~\cite{padovani14}. The integration of priorities into GV is adapted from Padovani and Novara~\cite{padovaninovara15}. Paraphrasing Dardha and Gay~\cite{dardhagay18extended}, priorities obey the following two laws:
\begin{enumerate*}[label= (\roman*) ]
\item an action with lower priority happens before an action with higher priority; and
\item communication requires \emph{equal} priorities for dual actions.
\end{enumerate*}

@@

Progress for the term language is standard for GV, and deviates from progress for linear $\lambda$-calculus only in that terms may reduce to values or \emph{ready terms}:
\begin{restatablelemma}{lempgvopenprogressterms}[Progress, $\tred$]%
  \label{lem:pgv-open-progress-terms}
  If $\tseq[\cs{p}]{\ty{\Gamma}}{M}{T}$ and $\ty{\Gamma}$ contains only session types, then:
  \begin{enumerate*}[label= (\alph*) ]
  \item $\tm{M}$ is a value;
  \item $\tm{M}\tred\tm{N}$ for some $\tm{N}$; or
  \item $\tm{M}$ is ready.
  \end{enumerate*}
\end{restatablelemma}
